\include{./preamble.tex}
\begin{document}

\maketitle

\section{CFMMs}
A new form of financial market, Decentralized Exchanges (DEXs) also referred
to as pool, have found its roots on top of blockchain technology. As in any
financial market, DEXs allow an agent to trade assets, in this case
cryptocurrencies (digital assets). Thorough a decentralized ledger stored on
multiple computers (blockchain) the DEXs keeping track of ownership by
storing in their storage slot. This report goes through the underlying
principles of DEXs and their properties by mathematical modeling and
optimization.

\subsection{Basics}
A DEX is governed by a constant function market maker (CFMM), which is a
function $\varphi:\mathbb{R}^{n}_{+} \to \mathbb{R}_+$ the trading function
and a set of conditions. Here $n \ge 2$ is the number of assets an agent can
trade, these are labeled $\{1,\ldots, n\}$. The CFMM is a function of the
reserves $y \in \mathbb{R}^{n}_+$, where $y_i \in \mathbb{R}_+$ is the
reserve of asset $i \in \{1,\ldots,n\}$, these are in the portfolio of the
DEX. There are two basic ways an agent can interact with the DEX:
\begin{itemize}
    \item Change liquidity
    \item Trade
\end{itemize}
This report covers the agents' iteration by trade. The liquidity change
could be added in the final master thesis, as part of a general introduction
the CFMMs.

An agent initializes a trade on a DEX, in which a set amount of assets is
traded for another. A set amount of assets given is denoted by $x \in
\mathbb{R}^{n}_+$, and the set amount of assets received by $\tilde{x} \in
\mathbb{R}^{n}_+$. Then $x_i, \tilde{x}_i \in \mathbb{R}_+$ is the amount of
given/received from asset $i \in \{1,\ldots,n\}$ respectively. A trade can either be
accepted or rejected, depending on an acceptance principle of the underlying
CFMM. If the trade is accepted the state of the pool is updated by changing
the reserves
\begin{align}
    y \mapsto y + x - \tilde{x}.
\end{align}
A necessary condition for trade acceptance is $y \ge 0$ after the trade. In
the case the trade is rejected the reserves do not change. Throughout we
also assume that the vectors $x$ and $\tilde{x}$ have disjoint support. This
is a valid assumption, since when trading, the agent is never giving and receiving a
the same asset, but rather one or multiple different ones. The
support of a vector $v \in \mathbb{R}^{n}_+$ is $\text{supp}(v) := \{i\; |\;
v_i > 0\}$, thus the assumption is $\text{supp}(x)\;\cap\;
\text{supp}(\tilde{x}) = \emptyset$.

A trade $(x, \tilde{x})$ is accepted, if the trading function $\varphi$
stays constant during the trade
\begin{align}
    \varphi(y+\gamma x - \tilde{x}) = \varphi(y).
\end{align}
Depending on the CFMM, a trade has a fee $\gamma \in (0, 1]$, where
$(1-\gamma)x$ of the set amount given assets is distributed to liquidity
providers for the reward of providing liquidity. The discounted amount
$\gamma x$ is actually being exchanged for $\tilde{x}$, but the reserves are
updated by the full amount $x - \tilde{x}$.

Throughout we assume that the trading function $\varphi$ is concave,
increasing and differentiable, as it is implemented by Uniswap
\cite{uniswap}, Balancer \cite{balancer}, Sushiswap \cite{sushiswap}, CurveFI
\cite{curvefi} and many others. Most of them are also homogenious
$\varphi(\alpha y)=\alpha(y)$, for all $\alpha \ge 0$.
\subsubsection{CFMMs Examples}

Weighted sum trading function
\begin{align}
    \varphi(y) = p^{T}y = \sum_{i=1}^{n} p_i y_i, \label{eq: weighted_sum}
\end{align}
where $p_i$ is called the price of asset $i$. The trade acceptance condition
simplifies to
\begin{align}
    &\varphi(y+\gamma x- \tilde{x}) = \varphi(y) \nonumber \\
    &p^{T} (y +\gamma x - \tilde{x}) = p^{T} y \nonumber \\
    &\gamma p^{T} x = p^{T} \tilde{x}.
\end{align}
The weighted geometric mean trading function
\begin{align}
    \varphi(y) = \prod_{i=1}^{n} y_i^{w_i}, \label{eq: weighted_geo_mean}
\end{align}
where $w_i > 0 $ and $\|w \|_1= 1$. We can simplify the trade acceptance
condition by splitting the product (ln of sum) in the indices of
$\text{supp}(x)$ and $\text{supp}(\tilde{x})$
\begin{align}
    \varphi(y+\gamma x- \tilde{x}) &= \varphi(y) \nonumber \\
    \prod\limits_{\substack{i \in \text{supp}(x) \\
    j \in \text{supp}(\tilde{x}) \\
    k \notin \text{supp}(x)\; \cup\; \text{supp}(\tilde{x})  }}
    (y_i + \gamma x_i)^{w_i}
    (y_j - \tilde{x}_j)^{w_j}
    (y_k)^{w_k} &= \prod_{i=1}^{n} (y_i)^{w_i} \nonumber\\
    \prod\limits_{\substack{i \in \text{supp}(x) \\
    j \in \text{supp}(\tilde{x})}}
    \left(1 + \frac{\gamma x_i}{y_i}\right)^{w_i}
    \left(1 - \frac{\tilde{x}_j}{y_j}\right)^{w_j}
    &= 1.\label{eq: wm_trade_acceptance}
\end{align}
For $|\text{supp}(x)| = |\text{supp}(\tilde{x})| = 1$, let $i$ and $j$ be the
corresponding indices of the giving and receiving asset then we can
explicitly derive an analytic expression for $x_i$ (or $\tilde{x}_j$) by
taking the power of $w_i$ (or $w_j$) and rearranging of equation
\ref{eq: wm_trade_acceptance}.
\begin{align}
&x_i = \frac{y_i}{\gamma}\left(
\frac{y_j^{w_j/w_i}}{\left(y_j - \tilde{x}_j \right)^{w_j/w_i}} - 1
\right)\\
&\tilde{x}_j= y_j\left(
    1 - \frac{y_i^{w_i/w_j}}{\left(y_i+\gamma x_i\right)^{w_i/w_j} }
\right)
\end{align}
Note that the weighted geometric mean is homogenious because the weights sum
up to one, i.e. $\| w \|_1 = 1$


Curve \cite{stableswap} market makers implement the following trading function
\begin{align}
    \varphi(y) =\|y\|_1 - \alpha \prod_{i=1}^{n} y_i^{-1} ,
\end{align}
for $\alpha > 0$, used for stable coin markets to reduce price impact without
large counter liquidity. A aggregated generalization is a convex combination between
the geometric weighted mean and the (un)weighted sum
\begin{align}
    \varphi(y) =(1-\alpha)\|y\|_1 - \alpha \prod_{i=1}^{n} y_i^{w_i},
\end{align}
where $\| w\|_1 = 1$, $w_i > 0$ and $\alpha \in [0,1]$. For $\alpha = 0$, the
trading function becomes the weighted sum and for $\alpha = 1$ the geometric
weighted mean.

\subsubsection{Concentrated Liquidity}
Another implementation of CFMM are automated market makers with concentrated
liquidity like UniswapV3 \cite{uniswapv3_paper}. The idea is to split
liquidity provision from the whole sublevel set
$\varphi(y +\gamma x-\tilde{x}) =\varphi(y)$ to multiple sets. The splitting
happens is determined by ticks $\{p(i)\}_{i \in \mathbb{Z} }$ and the rate
usually set $p(i) = 1.0001^{i}$, where $p(i)$ is the so called marginal
exchange rate of the tick. Every tick range
\begin{align}
    (p(i), p(i+1)],
\end{align}
has its own part of reserves (liquidity) and is governed by a one trading
function $\varphi$ depending on the liquidity bound by the tick range. The
trade is executed, when the marginal exchange rate is in the rage, on a
specifically provided liquidity. The liquidity providers specify an upper and
a lower bound $(p(U), p(L)]$ and provide liquidity on depth $\tilde{k} =
\varphi(y)$. They earn, split by weighted amount liquidity provided, the fees
of the trade associated with the CFMM.

During a trade when the marginal exchange rate raises above a trick range,
the tick range is flipped to the next one and the rest of the trade is
executed on the new tick range. This however is associated with more
transaction cost for the agent conducting the trade.
\subsection{First order approximations}
Let $\nabla \varphi(y) = P$ be the vector of unscaled prices, $P$ is positive
in all entries because $\varphi$ is by assumption increasing. The entries by
the gradient are
\begin{align}
    P_i = \frac{\partial \varphi(y)}{\partial y_i}
\end{align}
The reported price also called the marginal exchange rate is defined as
ration between the unscaled prices and the last unscaled price
\begin{align}
    p_i = \frac{P_i}{P_n},
\end{align}
where $p_n = 1$ always.

Recall the trade acceptance condition
\begin{align}
    \varphi(y + \gamma x - \tilde{x}) - \varphi(y) = 0
\end{align}
First order Taylor expansion of $(y + \gamma x - \tilde{x})$ around $y$ gives us
\begin{align}
    \varphi(y + \gamma x - \tilde{x}) - \varphi(y) &\simeq \nabla
    \varphi(y)^{T} (\gamma x - \tilde{x}) \\
                   &=P^{T} (\gamma x - \tilde{x}),
\end{align}
rearranging gives
\begin{align}
    \gamma \sum_{i \in \text{supp}(x) } P_i x_i \simeq \sum_{i \in
    \text{supp}(\tilde{x}) } P_i \tilde{x}_i,
\end{align}
which is scaling invariant (homogenious). Hence dividing by $P_n$
\begin{align}
    \gamma \sum_{i \in \text{supp}(x) } p_i x_i \simeq \sum_{i \in
    \text{supp}(\tilde{x}) } p_i \tilde{x}_i,
\end{align}
also holds. Note that this is a first order approximation, similarly to a
midpoint price in an order book. The actual trades are completely
determine by the trade acceptance condition.
\subsection{Swaps}
The most common type of trade interaction with a DEX is a swap, i.e. a two
asset trade. Denote these two assets with index $i$ and index $j$, then
$x=\delta e_i$, $\tilde{x}=\tilde{\delta} e_j$, where $\delta, \tilde{\delta}
\in \mathbb{R}_+$ and $e_i,e_j$ are the canonical basis vectors of
$\mathbb{R}^{n}$, $i \neq j$ (disjoint support). The trade acceptance
condition is then
\begin{align}
    \varphi(y + \gamma\delta e_i - \tilde{\delta}e_j) = \varphi(y).
\end{align}
It follows from the assumptions on $\varphi$, that the left hand side of the
equation $\varphi(y + \gamma\delta e_i - \tilde{\delta}e_j)$ is increasing
in $\delta$ and decreasing in $\tilde{\delta}$. Hence, for all $\delta \ge 0$
there is at most one $\tilde{\delta}$ which satisfies the trade acceptance
condition. Respectively for all $\tilde{\delta} \ge 0$ there is at most one
$\delta$ satisfying the trade acceptance condition. Thereby $\delta$ and
$\tilde{\delta}$ have a one to one function correspondence, and two asset
trades are uniquely determined by either $\delta$ or $\tilde{\delta}$.

In this case it is natural to define this function $F: \mathbb{R}_+ \to
\mathbb{R}$, and call it the forward exchange function. Where $F(\delta)$ is
unique to one $\tilde{\delta}$. The function $F$ is objective for all valid
trades, additionally the function is increasing since $\varphi$ is
increasing. Also $F$ is nonnegative, because $F(0) = 0$ and it is concave.
Use the implicit function theorem on the trade acceptance condition
\begin{align}
$f(\delta, \tilde{\delta}) = \varphi(y + \gamma\delta e_i -\tilde{\delta}e_j)
- \varphi(y) =0,
\end{align}
denote $y' = y+\gamma\delta e_i + \tilde{\delta}e_j$, the chain rule gives
\begin{align}
    F'(\delta) &= -\frac{\partial_\delta f(\delta,
    \tilde{\delta})}{\partial_{\tilde{\delta}} f(\delta, \tilde{\delta})}\\
       &=\gamma \frac{\left( \nabla_y \varphi(y') \right)_i}{\left( \nabla_y
       \varphi(y') \right)_j}
\end{align}
Together with the properties of $\varphi$, it can be shown that $F$ is
concave by showing that
\begin{align}
    F(\delta') \le F'(\delta)(\delta' - \delta) + F(\delta),
\end{align}
holds for all $\delta, \delta' \ge 0$ valid. Let $y'' = y + \gamma \delta' e_i -
F(\delta') e_j$ and note that $\varphi(y) = \varphi(y') = \varphi(y'')$ as
part of the trade acceptance condition. Since $\varphi$ is a concave function
\begin{align}
    \varphi(y'') \le \nabla_y\varphi(y')^{T} (y'' -y')+ \varphi(y'),
\end{align}
or simply
\begin{align}
    \nabla_y\varphi(y')^{T} (y'' - y') \ge 0.
\end{align}
Plugging in for $y''$ and $y'$ we get
\begin{align}
    &0 \le (\delta' -\delta)(\nabla_y \varphi(y'))_i - (F(\delta') -
    F(\delta))(\nabla_y \varphi(y'))_j \\
    &0 \le (\delta' -\delta)\frac{(\nabla_y \varphi(y'))_i}{(\nabla_y
    \varphi(y'))_j} - (F(\delta') - F(\delta))\\
    & F(\delta') \le F'(\delta)(\delta' - \delta) + F(\delta').
\end{align}

Respectively a definition of the reverse exchange function $G: \mathbb{R} _+
\to \mathbb{R} \cup {\infty}$, G(\tilde{\delta}) uniquely determined by
$\delta$ satisfying the trade acceptance condition. If no such $\delta$
exists then the function is mapped to $\{\infty\}$. The reverse exchange
function is nonnegative and increasing because of $\varphi$, but it is
convex. To show convexity, use the implicit function theorem
\begin{align}
    G'(\tilde{\delta})  &= -\frac{\partial_{\tilde{\delta}} f(\delta,
    \tilde{\delta})}{\partial_{\delta} f(\delta, \tilde{\delta})}\\
       &=\frac{1}{\gamma} \frac{\left( \nabla_y \varphi(y') \right)_j}{\left( \nabla_y
       \varphi(y') \right)_i}.
\end{align}
It needs to be shown that
\begin{align}
    G(\tilde{\delta}') \ge G'(\tilde{\delta})(\tilde{\delta}' -
    \tilde{\delta}) + G(\tilde{\delta})
\end{align}
Set $y'' = y + \gamma G(\tilde{\delta}')e_i - \tilde{\delta}'e_j$ similarly
with the forward exchange function we exploit the concavity of the trading
function $\varphi$ and the trade acceptance condition
\begin{align}
    &\nabla_y \varphi(y')^{T} (y'' - y') \ge 0\\
    &(G(\tilde{\delta}') - G(\tilde{\delta}))(\nabla_y \varphi(y'))_i
    - (\tilde{\delta}' - \tilde{\delta})(\nabla_y \varphi(y'))_j \ge 0\\
    &(G(\tilde{\delta}') - G(\tilde{\delta}))
    - (\tilde{\delta}' - \tilde{\delta})\frac{(\nabla_y
    \varphi(y'))_j}{(\nabla_y \varphi(y'))_i} \ge 0\\
    &G(\tilde{\delta}') \ge
    - (\tilde{\delta}' - \tilde{\delta})G'(\tilde{\delta}')
    + G(\tilde{\delta})
\end{align}




\section{General Trades}

\subsection{Utility optimization}

\subsection{Optimality Conditions}

\subsection{Arbitrage}

\subsection{Transaction Cost}

\nocite{bot_optimization}
\nocite{optimal_routing_cfmm}
\nocite{boyd_optimization}
\nocite{cfmms_basics}
\nocite{convex_network_flows}
\nocite{uniswap}
\nocite{balancer}
\printbibliography

\end{document}
