\include{./preamble.tex}


\begin{document}

\maketitle

Note:\newline
Formalization is garbage needs work.\newline
Correctness of the optimization problem/s is not tested.\newline


In decentralized finance (DeFi), on the blockchain, it is possible to exchange
an amount of one token for another one, essentially giving a token its value.
Generally these tokens are merely smart contracts, called ERC20 Smart
Contracts, thus the name ERC20 Tokens. A wallet on the blockhain "has" a
token, means that in the ERC20 smart contract of that token, there is stored
database entry, which is a key and a value. The key represents the wallets
address and the value is the amount of tokens the wallet has. Through
interaction with the tokens smart contract, one can among other things send
tokens to another address, or check how much tokens a specific wallet has.
Trading tokens works through Automated Market Makers (AMMs), so called token
pools, also smart contracts.
\newline

For example a pool which allows us to trade token X for token Y will have
liquidity, meaning it will contain an amount of $r_X$ , $r_Y$ of token X and
Y respectively, called reserves. A commonly used automated function market
maker is the constant product function $r_Xr_Y = k$, enforced to be constant
$k$ thought out a trade. If we are willing to trade $\Delta x$ of token X for
token Y, $\Delta x$ will be added to the reserve of token X and the amount of
token Y, $\Delta y$ we would get will be subtracted from the reserve of token
Y. Through the trade it holds that $(r_X + \Delta x)(r_Y - \Delta y) = k$.
And the amount $\Delta y$ of token $Y$ we would be getting can be calculated
implicitly from the this formula, i.e. $(r_X + \Delta x)(r_Y - \Delta y) =
r_Xr_Y$.
\newline

Example: $r_X = 100$, $r_Y = 50$, $\Delta x = 10$. In this case $r_Xr_Y = k =
5000$, and the amount $\Delta y = r_Y - \frac{r_Xr_Y}{r_X + \Delta x} \simeq
4.54$. Meaning we would be getting around $\Delta y = 4.54$ of tokens Y. The
updated reserves $r_X = 110$ and $r_Y = 44.46$.
\newline

An interesting thing for the constant product market maker is that an
infinite amount of tokens needs to be traded such that the reserve is
drained, i.e. $\Delta y = \lim_{\Delta x\to\infty} y - \frac{k}{x + \Delta x}
= y$. This enforces security of the pool and avoids pools being not usable.
\newline

In practice there is a protocol fee $1-\theta$ which is usually $0.3\%$. This
fee is collected from every trade and added to the liquidity, supplying the
liquidity providers. There is more things to explain here but not important
right now, later for so called Just in time liquidity. In short, instead of
trading the whole amount $\Delta x$ for $\Delta y$ as in the example above,
only $\theta \Delta x $ is traded. Meaning in practice the trade is constant
in the sence of $(r_X + \theta\Delta x)(r_Y - \Delta y) = r_Xr_Y$. And the
amount $\phi(\delta x) = \Delta y = r_Y - \frac{r_Xr_Y}{r_X + \theta \Delta
x}$ is received. Where $\phi: \mathbb{R}^{+} \to \mathbb{R}^{+}$ is the
trading function. Then the reserves are updated according to the trade.
Essentially the trade is conducted with $\Delta x$ with $\theta \Delta x$
only being traded but the whole amount $\Delta x$ added to the liquidity.
\newline

For now lets focus on the feeless regime, however I would like to explore the
calculations with the fee too.
\newline

The blockchain is constantly in motion each trade on a pool will alter the
tokens price, and a user committing a transaction to trade a token is
subjected to the quickness of the execution of the transaction to get a
desirable swap. A highly likely scenario is that a swap will happen before
the users transaction is executed (but committed), which will alter the
price such that the user will receive only a fraction of the token. Hence a
safety net is enforced called the slippage or slippage limit, which tells the
trade transaction how much of a price impact the user is willing to tolerate
to execute the trade. If the price change is more than the slippage limit,
the trade will not be executed and the transaction will be reverted. Lets say
a user commits a trade $\Delta x \in \mathbb{R}^{+}$ with slippage $\theta \in
(0,1)$, another user commits the trade first for $\Delta x_B$. The first
users trade will only be executed if $\phi(\Delta x_B + \Delta x) -
\phi(\Delta x_B) \ge (1-\theta)\phi(\Delta x)$.
\newline

Consider a scenario where a user (victim) has committed a transaction to
execute a trade, trading $\Delta x_V$ of Token X for Token Y with slippage
limit $\theta$. Each swap will alter the price, meaning that an attacker is
able commit a transaction to trade token X for token Y before the victim and
right after sell the amount back for token X, making a profit. The attacker
needs to maximize the amount $\Delta x_A$ tradable bound to the validity of
the trade of the victim, i.e.
\begin{align}
    \max &\quad \Delta x_A \\
    \text{s.t.:}&\quad \phi(\Delta x_V + \Delta x_A) - \phi(\Delta x_A) \ge
    (1-\theta) \phi(\Delta x_V)\\
        &\quad \Delta x_A \in \mathbb{R}^{+}
\end{align}
Q: What about with fees?\newline
Q: What about gas/transaction fee?\newline
Q: What about maximizing the profit?\newline
Q: What about a profit condition?\newline
Q: What about other CFMMs?
Q: Slippage amount is not easily extractable from the information provided in
the transaction data, how to get the correct slippage amount?
Q: Propose a CFMM or a smart contract preventing this attack?
\newline

Next consider a arbitrage scenario. Tokens X and Y and two pools with
reserves $r_{i, j}$, $i \in \{1, 2\}$ and $j\in \{X, Y\}$. In the constant
motion of the blockchain the two pools may become in a state of unbalance,
making the prices of the pools unequal. An arbitrageur exploiting this
scenario will buy/sell from one pool and then sell/buy from the other pool.
Let us say that the arbitrageur has detected that that the optimal trade is
$X \to_{\text{Pool 1}} Y \to_{\text{Pool 2}} X$. Q: How to detect the
arbitrage opportunity?. What is the maximal amount $\Delta x$ he can trade to
get $\Delta x'$ out, maximizing the profit $\Delta x' - \Delta x$? The trades
need to be subjected to the constant function market maker, essentially the
constant product needs to be constant through out the trade in pool one and
then through the trade in pool 2.
\begin{align}
    \max &\quad \Delta x' - \Delta x \\
    \text{s.t.:}&\quad(r_{1,X} + \Delta x)(r_{1,Y} - \Delta y) =
    r_{1,X}r_{1,Y} \\
    &\quad(r_{2,X} - \Delta x')(r_{2,Y} + \Delta y) = r_{2,X}r_{2,Y},
\end{align}
where $\Delta y = r_{1,Y} - \frac{r_{1,X}r_{1,Y}}{r_{1,X} + \Delta
x}$.\newline
Q: How to find the arbitrage opportunity ? Graphs ? \newline
Q: Is there a profit condition or only brute force ? \newline
Q: Optimization with multiple pools $X\to Y \to Z \to \ldots \to X$ ? \newline
Q: What is the optimal routing ? \newline
Q: Other CFMMs ? \newline
Q: With fees ? \newline
Q: With transaction/gas fees ?

On simple pools liquidity providers are bound to provide liquidity on the
whole price curve $r_Xr_Y = k$. A newly adopted mechanic is to allow
liquidity providers to provide liquidity on a specific price range. This
makes the price impact after a trade different depending on how much
liquidity is provided on this specific price range. Consider a scenario where
a user (victim) is executing a large trade, an attacker sees this trade
before its executed in the transaction pool and decides to provide liquidity
before the trade of the victim is executed, and after the trade collecting it
earning himself the majority of the fees.\newline

Q: What is the optimal amount of liquidity to provide, formalization? \newline
Q: Is there an upper bound on the profit, depending on the other liquidity
providers ?






\end{document}
