\include{./preamble.tex}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
% Node styles
\tikzstyle{new style 0}=[fill={rgb,255: red,128; green,128; blue,128},
draw=black, shape=circle, inner sep=0.5mm]
\tikzstyle{1}=[fill=white, draw=black, shape=circle, font=\scriptsize,
        inner sep=0.5mm, outer sep=0.5mm, minimum width=0.5mm]

% Edge styles
\tikzstyle{new edge style 0}=[-, draw={rgb,255: red,191; green,191; blue,191}, ultra thin]
\tikzstyle{new edge style 1}=[-, draw={rgb,255: red,191; green,0; blue,64},
ultra thick]
\tikzstyle{new edge style 2}=[-, draw={rgb,255: red,191; green,0; blue,64}, in=60, out=120, dotted]

\begin{document}

\maketitle

\section{Learning Setup}
This report will go through the RL framework \& neural network model used
find profitable walks in the arbitrage problem. Firstly the agent and the
environment for the setting is defined together with objective and the
rewards. Secondly Graph Neural Network and the training framework are
outlined. The distinct case here is that model is trained on the line graph
of the defined problem graph. The reason for this choice is for
generalization purposes. The last section explains the rough process of
obtaining and filtering the data necessary as an input for the problem and
learning.

\subsection{Agent and Environment}
The framework (or the game) the agent is trying to learn is finding a walk
in a graph that does not use repeated edges (trail), starts and ends at
the same node (circular) and is profitable. The graph is represented by ERC20 \cite{erc20}
tokens as edges and decentralized exchange pools as edges. Each pool has a
swap fee, which usually is in the range of $0.3\%-1\%$, and a marginal
exchange rate. The marginal exchange rate, unique to the pool, is inversely
bidirectional. Meaning, while swapping token $0$ for token $1$ in the pool
the usual marginal exchange rate is used, but when swapping from token $1$ to
token $0$ the inverse marginal exchange rate is used. In both cases the
referred term is marginal exchange rate. `Used' here is meant in the context
of the learning setup, in the actual setting the automated market maker
formula is used to determine the amount exchanged.
\newline

The agent, starting from the given node (token) chooses actions from a set of
valid actions. Actions are moves to the next node via an edge (pool). The
valid actions rule out the edges that have been used already. The idea behind
this is the following. If there is an arbitrage opportunity, and the trades
are conducted in a chronological order, at each swap the pool's `capacity'
will be exerted and the pool's exchange rate will in balance with the overall
market. Hence, the pool has no liquidity for an arbitrage to be used again.
\newline

When the agent arrives at a terminal state, there are two possibilities.
Firstly the agent has arrived back to its starting point, where the product
of the exchange rates is calculated to give the profit of the walk. In this
case, at each trade a trading penalty is utilized to account for the fees
of each transaction. The second case is where the agent arrives at a dead
end, meaning the agent has not reached the starting point and has no valid
actions left.

\subsubsection{Rewards}
When reaching a terminal state the agent is rewarded a different amount. In
the case the agent is at a dead end the reward is
\begin{align}
    r_L = -1.
\end{align}
In the case the agent ends back up at the origin, the reward is the sum of
the natural logarithm of the marginal exchange rates $Z_p$ in the path $P$
and the trade penalty $\tau$. Additionally, the sum is adjusted by a
win-loss multiplier $M$ to give the reward
\begin{align}
    r_W = M \cdot \sum_{p \in P} \log(Z_p \cdot \tau)
\end{align}
The idea behind using the sum of the natural logarithm of the marginal
exchange rates is that it accounts to the sequence of trades that give no
profit. A sequence of trades will be a profitable if the product of the
marginal exchange rates is bigger than $1$ and a not profitable if it is
smaller. The logarithm will be positive for a profitable trade and negative
for a nonprofitable one. Another reason for using the logarithm is to adjust
for the high discrepancy of the marginal exchange rates between the pools.
\newline

The use of the trade penalty $\tau$ is generally a good idea since it
accustoms for the gas fee of each transaction. But it might prevent the
algorithm from finding any profitable walks because they are actually rare,
since the market is usually in balance. Thus, in the training process the use
of this parameter is omitted by setting $\tau = 1$. In actual production it
is useful to set $\tau \in [0.95, 1)$, or even an adaptive value. The value
of the win-loss multiplier is set to $M=2$ for training.

\subsection{Deep Graph Neural Network}
The idea is to train a model that is general in the sense that it can be used
on smaller or bigger graphs, graphs of different node and edge size. Encoding
the state and pushing it through a traditional Deep NN will give us the same
policy dimension every time, which depends on the architecture of the
network. Instead, the aim is to always get a policy vector that has the size
of two times the amount of pools, since the agent can move in both directions
of the pool (inverse marginal exchange rate). This is essentially the same
amount as the edges in the problem graph. A good idea is to use a graph
neural network, but the output is a policy vector of the size of the number of
nodes in the graph, which have no connection to the number of edges in the
graph. To make the model more general the model will be trained on a graph
neural network based on the line graph of the problem graph. The line graph of a
graph has node for each edge in the actual graph. Two nodes are connected
by an edge if and only if their edges in the original graph share a common
endpoint. Each node (pool) in the line graph will have five attributes:
\begin{itemize}
    \item Marginal exchange rate
    \item Pool swap fee
    \item Binary entry if used or not
    \item Binary entry if the agent is currently at token 0 of this pool
    \item Binary entry if the agent is currently at token 1 of this pool
\end{itemize}
The encoder computes the node embeddings $h_i^{0}$ through a learned linear
projection to dimension $d_h = 320$. The embeddings are then updated via. $N$
Residual Multihead Attention Layers \cite{gatv2}. Rather than the one from
the original paper \cite{gan_velickovic}, an updated formula for the
attention weights is used. Each layer consists of two sublayers, a multihead
attention layer (GAT) with $20$ heads, for message passing between the nodes
which is flattened by a linear connection back to the original dimension
$d_h$. And a node-wise fully connected feed-forward layer (FF) with $1280$
hidden dimension. Both sublayers have a Batch Norm (BN) and a residual
connection similar to \cite{rl_coopt}. The following is a sketch of the
residual block
\begin{align}
    \hat{h}_{i} &= \text{BN}^{l}\left( h_i^{(l-1)}
    + \text{GAT}_i^{l}(h_1^{(l-1)},\ldots,h_n^{(l-1)})\right)\\
    h_i^{(l)} &= \text{BN}^{l}\left( \hat{h}_{i} + \text{FF}_i^{l}(\hat{h}_i) \right).
\end{align}
The decoder has two heads, the policy head and the value head as in
\cite{alphazero}. The policy head has a graph Multihead attention layer with
1 head with Batch Norm forwarded to a linear layer, which expands the output
per node by two with a Relu activation function. The value head has also a
graph Multihead attention layer with 1 head fed through a linear layer and
taken the mean. More details will be provided upon full finishing the project
and settling with a finetuned architecture.
\subsection{Training}
The training is purely unsupervised based on the MCTS algorithm. Firstly the
agent `plays' the network from a random state (node \& marginal exchange
rates) this is called the self play stage. At each self play The marginal
exchange rates of the prices are chosen randomly, all from a random block
number $t$ on the Ethereum blockchain. The next state is chosen based on
$N_{\text{MCTS}}$ searches. In each iteration the agent plays the network
$N_{\text{play} }$ times until reaching a terminal state. Upon reaching the
terminal state, the state, the policy vector and the values of the MCTS
algorithm along the state path are saved. After the self play stage the model
parameters are updated based on the self play data with batch number
$N_{\text{batch} }$ and $N_{\text{epoch} }$ epochs. The loss function
consists of separate policy loss, a cross entropy loss with mean
reduction and a value loss, a mean-squared loss. These are then added
together for a full loss function.
\begin{align}
    L(\theta) = L_{\text{policy}}(\theta) + L_{\text{value}}(\theta),
\end{align}
where $\theta$ are the parameters of the model and
\begin{align}
    L_{\text{policy}} (\theta) &= - \sum_{i=1}^{N_{\text{batch}}}
    \pi(a^{(i)}_{t_i} |s ^{(i)}_{t_i})
    \ln \left(\text{softmax}\left(\tilde{\pi}(a^{(i)}_{t_i}|s^{(i)}_{t_i};
        \theta\right)\right) \\
    L_{\text{value}} (\theta) &= \frac{1}{N_{\text{batch}}}
    \sum_{i=1}^{N_{\text{batch}}}
    \left( v(s^{(i)}_{t_i}) - \tilde{v}(s^{(i)}_{t_i}; \theta)  \right)^{2}.
\end{align}
The weighs of the model are updated using the Adam optimizer \cite{adam} with
a learning rate of $\gamma = 10^{-3} $, a weight decay of $\lambda =
10^{-4}$, forgetting factors $\beta_1 = 0.9$, $\beta_2 = 0.999$ and
$\varepsilon = 10^{-8}$ to prevent division by zero.
\newline

For now $N_{\text{batch} }=128$, $N_{\text{epochs}} = 10$, $N_{\text{play}} =
100$, $N_{\text{MCTS}} = 50$ with $100$ iterations. The marginal exchange
rates are pulled from the block range $t \in [18\cdot10^{6}, 21.8\cdot10^6]$.
The number of residual block layers is $10$, these do not share weights. The
UCB1 exploration parameter $C$ in the MCTS algorithm for choosing the next
child in the tree is initialized as $C=3.5$ for the first third of the
iterations for the second third $C=2.0$ and for the last third $C=1.5$.
\subsection{Data}
The data used is publicly available on the Ethereum mainnet blockchain and
can be accessed through an archive node query. The main data source are all
the UniswapV2 and UniswapV3 \cite{uniswap}. The difference between V2 pools
and V3 pools is that the V3 pools have concentrated liquidity pools, but both
the follow constant product function. At the current time of writing there
are around $4.3\cdot10^{5}$ pools and approximately $4.1\cdot10^{5}$
different ERC20 \cite{erc20} tokens in these pools. To narrow down the
meaningful pools, firstly we use pools that have been created before block
$18\cdot10^{6}$ and secondly we use a degree filter. . Tokens which have a
smaller degree then $15$ are filtered out. Left are $1885$ pools and $150$
different tokens. The question still remains if these pools are still in use
currently. Hence, the pools with low liquidity are also filtered out, by
looking at their two token reserves, the $\text{ETH}$ value of the
reserves is crucial here. If the reserve is worth less then $100\; \text{ETH} $
it is filtered out. Left are $127$ pools and $66$ tokens. For the training
process the marginal exchange rates of the pools are extracted from block
$18\cdot10^{6}$ until the current block every $9000$ blocks which is equal
to around every $30$ hours ($\sim12s$ is the time until the next block). The
prices are then randomly sampled for training.

\section{Code}
The code can be checked on my Github \cite{github}.

\nocite{github}
\nocite{gatv2}
\nocite{gan_velickovic}
\printbibliography

\end{document}
