\addchap*{Abstract}
Decentralized Exchanges (DEXs) are a new form of financial markets emerging
on top of blockchain technology. Through these arises a competitive and
complex environment of Maximal Extractable Value (MEV), where entities called
Searchers race against the block time and compete against each other to find
profitable opportunities when the market is at rest, and the next block is
not yet committed. One of these opportunities is an $n$-cyclic arbitrage, on
which this thesis focuses on. Through modeling of the building blocks of
DEXs, governed by Constant Function Market Makers (CFMMs), we can derive a
nonlinear convex optimization problem finding an optimal $n$-cyclic arbitrage
trade. However this optimization problem neither specifies the order in which
these trades should be executed, nor reduces optimization to specific, used
DEXs. To combat this, we propose a Markov Decision Process and find an
optimal policy through a Monte Carlo Tree Search REINFORCE algorithm guiding
us to a profitable $n$-cyclic arbitrage walk based on the marginal exchange
rates. The information of the found walk can be then applied to the nonlinear
optimization problem to find an optimal trade in a timely manner.

\addchap*{Vorwort}
Dezentrale Börsen (DEXs) sind eine neue Form von Finanzmärkten, die auf der
Grundlage der Blockchaintechnologie entstehen. Durch sie entsteht ein
wettbewerbsintensives und komplexes Umfeld des Maximal Extractable Value
(MEV), in dem Akteure, sogenannte Searchers, gegen die Blockzeit antreten und
miteinander konkurrieren um profitable Gelegenheiten zu finden, während der
Markt inaktiv ist und der nächste Block noch nicht bestätigt wurde. Eine
dieser Gelegenheiten ist eine n-zyklische Arbitrage, auf die sich diese
Arbeit konzentriert. Durch die Modellierung der Grundbausteine von DEXs,
den Constant Function Market Makers (CFMMs), lässt sich ein nichtlineares
convexes Optimierungsproblem ableiten, das einen optimalen n-zyklischen
Arbitrage Trade identifiziert. Dieses Optimierungsproblem gibt jedoch weder
die Ausführungsreihenfolge der Trades vor noch beschränkt es die
Optimierung auf konkret genutzte DEXs. Um dies zu adressieren, schlagen wir
ein Markow Entscheidungsproblem vor und ermitteln eine optimale Strategie
mithilfe eines Monte Carlo Baum Suche REINFORCE Algorithmus, der uns anhand
der nominellen Wechselkurse zu einem profitablen n-zyklischen Arbitragepfad
führt. Die Informationen des gefundenen Pfads können anschließend in das
nichtlineare Optimierungsproblem genutzt werden, um zeitnah einen optimalen
Trade zu bestimmen.
