\chapter{Background Optimization}
In this chapter we will go through the necessary background to understand
optimality conditions for nonlinear constrained optimization problems. We
will introduce necessary optimality conditions. In the case we have a \"nice
enough\", convex, set we can show that these condition become sufficient for
optimality. These are all very well known results from the theory of
nonlinear optimization, hence this chapter is heavily based on the lecture
notes \cite{bot_optimization}.
\section{First Order Optimality Conditions}
Generally we are considering a nonempty subset $X \subseteq \mathbb{R}^{n}$
for $n \in \mathbb{N}$ and a function $f: \mathbb{R}^{n}  \to \mathbb{R}$ in
the context of an optimization problem
\begin{align}
    \min_{x \in X}\; &f(x) \label{eq: basic_opt}.
\end{align}
A local minimum of this problem is an element $x^{*} \in X$,
such that there is an $\varepsilon > 0$ ball, wrt. to the euclidean norm,
i.e.
\begin{align}
    U_\varepsilon(x^{*}) = \{x \in X: \|x - x^{*}\| < \varepsilon\},
\end{align}
such that
\begin{align}
    f(x^{*} ) \le f(x) \qquad \forall x \in U_\varepsilon(x^{*})\cap X.
\end{align}
A global minimum is an element $x^{*} \in X$ such that the inequality holds
on the whole set $X$.
\begin{align}
    f(x^{*}) \le f(x) \qquad \forall x \in X.
\end{align}
To better understand local minima we can introduce look at directions, from a
given point $x_0$, from where we can move and still remain in $X$. In this
sense we introduce a cone
\begin{align}
    T_X(x_0) := \left\{ d \in \mathbb{R}^{n}\Big|\;\exists (x^k)_{k \in
    \mathbb{N}} \subseteq X,\; \exists (t_k)_{k \in \mathbb{N} }\searrow
0\;\; : \lim_{k \to \infty} \frac{x^k-x_0}{t_k} = d  \right\},
\end{align}
called the Bouligard tangent cone. It is nonempty since $0 \in T_X(x_0)$ for
the choice of $x^{k} = x_0$ and $t_k = \frac{1}{k}$. It is really a cone
since it is scaling invariant for $d \in T_X(x_0)$ also $\lambda d \in
T_X(x_0)$, for all $\lambda > 0$, by multiplying $(t_k)_{k \in \mathbb{N} }$
with $\frac{1}{\lambda}$.
\begin{proposition}
    \label{prop: descent}
    let $x^{*}$ be a local minimum of \ref{eq: basic_opt}, where $f \in
    C^1(U_{\varepsilon}(x^{*}))$, for an $\varepsilon >0$. Then
    \begin{align}
        \nabla f(x^{*})^{T}d \ge 0 \qquad \forall d \in T_X(x^{*}).
    \end{align}
\end{proposition}
\begin{proof}
Fix $d \in T_X(x^{*})$, then there are sequences $(x^{k})_{k \in \mathbb{N} }$
and $(t_k)_{k \in \mathbb{N}}$ such that
\begin{align}
    \lim_{k \to \infty} \frac{x^{k}-x^{*}}{t_k} = d.
\end{align}
Multiplying by $t_k$ gives
\begin{align}
   \lim_{k \to \infty} x^{k} - x^{*} =
   \lim_{k \to \infty}  \frac{x^{k}-x^{*}  }{t_k}t_k =
   \lim_{k \to \infty} dt_k = 0.
\end{align}
Since $x^{*}$ is a local minimum of \ref{eq: basic_opt} we have that
$f(x^{*}) \le f(x)$ for all $x \in U_\varepsilon(x^{*}) \cap X$. So there is
a $k_\varepsilon \in \mathbb{N}$ such that $x^{k} \in U_\varepsilon(x^{*})$
for all $k\ge k_\varepsilon$. By the mean value theorem there is a $\xi^{k}
\in \{\lambda x^{*} + (1-\lambda)x^{k}| \lambda \in [0, 1]  \} $, such that
\begin{align}
    f(x^{k}) - f(x^{*} ) = \nabla f(\xi^{k} )(x^{k} -x^{*} ).
\end{align}
Let $\xi^{k} = \lambda_k x^{*} + (1-\lambda_k)x^{k}$, then for all $k \ge
k_\varepsilon$ we have
\begin{align}
    \|\xi^{k} -x^{k} \| &= \|\lambda_k x^{*} + (1-\lambda_k) x^{k}
    -x^{*}\| \\
                &=|\lambda_k-1| \| x^{k}  - x^{*} \|\\
                &\le \| x^{k}  - x^{*} \| \to 0.
\end{align}
And thus $\xi^{k} \to x^{*}$. By the continuity of the gradient in the
neighborhood of $x^{*}$ we also have $\nabla f(\xi^{k}) \to \nabla f(x^{*})$,
thereby
\begin{align}
    \nabla f(x^{*})^{T} d &= \lim_{k \to \infty} \frac{\nabla f(\xi^{*})
    (x^{k}-x^{*})}{t_k} \\
              &=\lim_{k \to \infty} \frac{f(x^{k}) - f(x^{*})  }{t_{k} } \\
              &\ge 0.
\end{align}
\end{proof}
Basically statement in Proposition \ref{prop: descent} says that at a local
minimum $x^{*}$ there is no tangent direction or \"descent direction\" for
which $\phi '(0) < 0$, where $\phi(t) := f(x^{*} + t d)$.

Bringing everything together give the first order optimality condition.
\begin{theorem}
    Let $x^{*}$ be a local minimum of \ref{eq: basic_opt} and $f \in C^{1}
    (U_\varepsilon(x^{*} )$ for an $\varepsilon >0$. Then $x^{*}$ is a
    critical point of $f$, i.e.
    \begin{align}
        \nabla f(x^{*}) = 0.
    \end{align}
\end{theorem}
\begin{proof}
    Since $x^{*}$ is a local minimum of \ref{eq: basic_opt} and $f \in C^{1}
    (U_\varepsilon(x^{*} ))$ then $x^{*} $ is also a local minimum in of f in
    $U_\varepsilon(x^{*})$. By Proposition \ref{prop: descent} we have that
    $\nabla f(x^{*} )^{T}  d \ge 0$, for all $d \in T_{U_\varepsilon(x^{*})
    }(x^{*} ) = \mathbb{R} ^{n}$. Fix $d \neq 0$ and plug in $d$ and $-d$,
    giving $\nabla f(x^{*})^{T} d = 0$. Hence $\nabla f(x^{*} ) = 0$.
\end{proof}
From now on consider a particular set $X$, the feasible set, defining a
constrained optimization problem through
\begin{align}
    X = \left\{ x \in \mathbb{R}^{n}\Big|
        \begin{array}{l}
            g_i(x) \le 0,\; i=1,\ldots,m\\
            h_j(x) = 0,\; j=1,\ldots,p
    \end{array} \right\}.
\end{align}
Then we have a general optimization problem
\begin{align}
    \min\;&f(x) \label{eq: gen_constraned}\\
    \text{s.t.:}\;&g_i(x) \le 0,\; i=1,\ldots,m \nonumber\\
        &h_j(x) = 0,\; j=1,\ldots,p \nonumber\\
        &x \in \mathbb{R}^{n}\nonumber,
\end{align}
i.e. $x \in X$, in the feasible set.

Generally the Bouligard tangent cone $T_X(x_0)$ for a $x_0 \in X$ is not a
practical object, since it cannot be determined out of the box. Hence we can
introduce the linearized tangent cone, which will serve as a replacement and
can be easily determined for a given feasible set $X$ and in a specific
setting will be valid representation of the Bouligard tangent cone. For an
element $x_0 \in X$ consider
\begin{align}
    \mathcal{A}(x_0) = \left\{ i = 1,\ldots,m|\;g(x_0) = 0 \right\},
\end{align}
the set of active indices at $x_0$ and
\begin{align}
    \mathcal{I}(x_0) = \left\{1,\ldots,m \right\}\setminus\mathcal{A}(x_0).
\end{align}
Then the linearized tangent cone of $X$ at $x_0$ is defined as
\begin{align}
    T_{\text{lin} }(x_0) = \left\{
        d\in \mathbb{R}^{n} \Big|
        \begin{array}{l}
            \nabla g_i(x_0)^{T} d \le 0,\; i \in \mathcal{A}(x_0)\\
            \nabla h_j(x_0)^{T} d = 0,\; j=1,\ldots,p.
        \end{array} \right\}.
\end{align}
The linearized tangent cone is really a cone since the scaling with
$\lambda>0$ does not change the outcome of the equalities and inequalities.
\begin{lemma}
    Let $x_{0} \in X$, then it holds that $T_X(x_0) \subseteq
    T_{\text{lin}}(x_0)$.
\end{lemma}
\begin{proof}
    We start with a $x_0 \in X$ and $d \in T_X(x_0)$. Then there are
    sequences from the Bouligard tangent cone $(x^{k})_{k\in N}$ and
    $(t_k)_{k \in N} \searrow 0 $, such that
    \begin{align}
        \lim_{ k \to \infty} \frac{x^k-x_0}{t_k} = 0.
    \end{align}
    First we prove that $\nabla g_i(x_0)^{T} d \le 0$ for all active indices
    $i \in \mathcal{A}(x_0)$. For this consider the mean value theorem, for
    all $k\ge1$ there is a $\xi^{k}  \in \{\lambda x_0 + (1-\lambda) x^{k}|\;
    \lambda \in [0, 1]\} $ such that
    \begin{align}
        g_i(x^{k}) - g_i(x_0) = \nabla g_i(\xi^{k})^{T} (x^{k} - x_0).
    \end{align}
    Since $i$ is active then we have $g_i(x_0) = 0$ and thus
    \begin{align}
         \nabla g_i(\xi^{k})^{T} (x^{k} - x_0) = g_i(x^{k}) \le 0.
    \end{align}
    Dividing by $t_k$ and passing the limit gives
    \begin{align}
        0 \ge \lim_{k \to \infty} \nabla g_i(\xi^{k})^{T}  \frac{x^k-x_0}{t_k}
        = \nabla g_i(x_0)^{T} d.
    \end{align}
    For the second statement in $T_{\text{lin}}(x_0)$ we need to show that
    $\nabla h_j(x_0)^{T} d = 0$ for all $j=1,\ldots,p$. For this we can use
    the same argumentation, by the mean value theorem for
    all $k\ge1$ there is a $\xi^{k}  \in \{\lambda x_0 + (1-\lambda) x^{k}|\;
    \lambda \in [0, 1]\} $ such that
    \begin{align}
        h_j(x^{k}) - h_j(x_0) = \nabla h_j(\xi^{k})^{T} (x^{k} - x_0).
    \end{align}
    Since $h_j(x_0) = 0$ and $h_j(x^{k}) = 0$  we have that
    \begin{align}
        \nabla h_j(\xi^{k})^{T} (x^{k} - x_0) = 0.
    \end{align}
    Dividing by $t_k$ and passing the limit gives
    \begin{align}
        0 = \lim_{ k \to \infty} \nabla h_j(\xi^{k})^{T} \frac{x^{k} - x_0}
        {t_k}= \nabla h_j(x_0)^{T} d.
    \end{align}
    Thus $d \in T_{\text{lin}}(x_0)$.
\end{proof}

\begin{lemma}[Lemma of Ferkas]
    \label{lem: ferkas}
    Let $A \in \mathbb{R}^{m \times n}$ and $b \in \mathbb{R}^{m}$. Then
    the two statements are equivalent
    \begin{enumerate}
        \item The system $Ax = b$ has a solution $x\ge0$.
        \item For all $d \in \mathbb{R}^{m}$ with $A^{T} d \ge 0$ it holds
            that $b^{T} d  \ge 0$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    We can use the strong duality result to prove equivalence.
    For (1) $\Rightarrow$ (2), consider
    \begin{align}
        \text{min}\; & 0^{T} x\\
        \text{s.t.:}\;& Ax = b \nonumber \\
            & x \ge 0. \nonumber
    \end{align}
    Thus the dual of wrt. $d \in \mathbb{R}^{m}$ is
    \begin{align}
        \text{max}\; & b^{T}d \\
        \text{s.t.:}\;& Ad \ge 0 \nonumber
    \end{align}
    By strong duality the optimal value of the dual is $0 \in
    \mathbb{R}^{m}$. Hence, for all $d$ with $A^{T}d \ge 0$ we have that
    $b^{T}d \ge 0$.

    For (2) $\Rightarrow$ (1), we have that $b^{T} d \ge 0$ for all $d$ with
    $A^{T} d \ge 0$. By strong duality the optimal value of the primal is
    $0$, and by the dual the optimal solution needs to be $x\ge 0$. By
    feasibility of the primal there is an $x\ge0$ satisfying $Ax=b$.
\end{proof}
\begin{theorem}
    \label{thm: duality}
    Let $x_0 \in X$, the it holds that
    \begin{align}
        -(T_{lin}(x_0))^{*} = N_{\text{lin} }(x_0).
    \end{align}
    where
    \begin{align}
        N_{\text{lin}} =
        \left\{
            \sum_{i=1}^{m} \lambda_i \nabla g_i(x_0)
            + \sum_{j=1}^{p} \mu_j\nabla h_j(x_0)
             \left|
             \begin{array}{l}
                \lambda_i \ge 0,\; i \in \mathcal{A}(x_0),\\
                \lambda_i = 0,\; i \in \mathcal{I}(x_0),\\
                \mu_j \in \mathbb{R},\; j=1,\ldots,p
            \end{array}
        \right\} \right.
    \end{align}
\end{theorem}
\begin{proof}
    Direction $\supseteq$: Let $s \in N_{\text{lin} }(x_0)$ and $d \in
    T_{\text{lin} }(x_0)$. Then
    \begin{align}
        -s^{T} d &= -\sum_{i=1}^{m} \lambda_i\nabla g_i(x_0)^{T}d
        -\sum_{j=1}^{p} \mu_j \nabla h_j(x_0)^{T} d\\
                 &\ge 0.
    \end{align}
    Because $\lambda_i \ge0$, $\nabla g_i(x_0)^T d \le 0$ and $\nabla
    h_j(x_0)^{T} d =0 $. Hence $s \in -(T_{\text{lin} }(x_0))^{*}$.

    Direction $\subseteq$: Let $s \in -(T_{\text{lin} }(x_0))^{*}$, then by
    definition $-s^{T} d \ge 0$ for all $d \in T_{\text{lin}}(x_0)$. We can
    use Lemma \ref{lem: ferkas} with the matrix $A \in \mathbb{R}^{n\times
    (r + 2p)}$ with $\mathcal{A}(x_0) = \{i_1, \ldots, i_r\} $ and
    \begin{align}
        A:=\begin{pmatrix} A_g, A_h, -A_h
        \end{pmatrix},
    \end{align}
    where
    \begin{align}
        A_g &=
        \begin{pmatrix}
        -\nabla g_{i_1}(x_0),\ldots,-\nabla g_{i_r}(x_0)
        \end{pmatrix},\\
        A_h &=
        \begin{pmatrix}
            \nabla h_1(x_0),\ldots,\nabla h_p(x_0)
        \end{pmatrix} .
    \end{align}
    By Lemma \ref{lem: ferkas} there is a $\nu = (\lambda, \mu_1, \mu_2) \in
    \mathbb{R}_+^{r}\times \mathbb{R}^{p}_+\times \mathbb{R}_+^{p}$
    satisfying
    \begin{align}
        A\nu = -s,
    \end{align}
    which is equivalent to
    \begin{align}
        \sum_{i=1}^{m} -\lambda_i\nabla g_i(x_0)
        +\sum_{j=1}^{p} (\mu_1)_j\nabla h_j(x_0)
        +\sum_{j=1}^{p} -(\mu_2)_j\nabla h_j(x_0) = -s.
    \end{align}
    By setting $\mu = \mu_2 - \mu_1$ and the fact that $\lambda_i \ge 0$ for
    all $i \in \mathcal{A}(x_0)$ and $ \lambda_i =0$ for all $i \in
    \mathcal{I}(x_0)$ we have
    \begin{align}
       s =  \sum_{i=1}^{m} \lambda_i\nabla g_i(x_0)
        +\sum_{j=1}^{p} \mu_j\nabla h_j(x_0),
    \end{align}
    thus $s \in N_{\text{lin}}(x_0)$.
\end{proof}

Next we will introduce the KKT conditions
\begin{definition}
    The function $L: \mathbb{R}^{n} \times \mathbb{R}^{m} \times
    \mathbb{R}^{p} \to \mathbb{R}$ defined as
    \begin{align}
        L(x, \lambda, \mu)= f(x) + \lambda^{T} g(x) + \mu^{T} h(x)
    \end{align}
    is called the Lagrangian function of the optimization problems
    \ref{eq: gen_constraned}. The conditions
    \begin{align}
        \begin{cases}
            \nabla_x L(x, \lambda, \mu) = 0,\\
            \lambda \ge 0,\; g(x) \le 0,\; \lambda^{T} g(x) = 0\\
            h(x) = 0
        \end{cases}
    \end{align}
    are called KKT optimality conditions, where
    \begin{align}
        \nabla_x L(x, \lambda, \mu) &= \nabla f(x) + \lambda^{T} \nabla
            g(x) +\mu^{T} \nabla h(x)\\
        &= \nabla f(x) + \sum_{i=1}^{m} \lambda_i \nabla g_i(x)
            + \sum_{j=1}^{p} \mu_j \nabla h_j(x).
    \end{align}
    An element $(x^{*}, \lambda^{*}, \mu^{*})$ fulfilling  the KKT optimality conditions is called a KKT-point. The vectors $\lambda^{*}$ and $\mu^{*}$ are
    called Lagrange multipliers associated with the corresponding restrictions,
    i.e. $g(x^{*}) \le 0$ and $h(x^{*}) = 0$.
\end{definition}
In the case there are no restrictions the KKT optimality conditions are
$\nabla f(x) = 0$. Thus the interpretation of these conditions is a
generalization of the first order optimality for constrained optimization and
these play are necessary conditions for a solution to be optimal. The KKT
conditions are sufficient for a local optimum, however in the case the
problem is convex the conditions are sufficient.
\begin{definition}
    Let $x_0 \in X$, where $X$ is the feasible set from equation \ref{eq:
    gen_constraned}. Then $x_0$ fulfills Abadie constraint qualification (CQ)
    if
    \begin{align}
        T_X(x_0) = T_{\text{lin}}(x_0).
    \end{align}
\end{definition}
A particularly nice thing about this CQ is that it does not assume anything
about the objective function $f$ and gives a very nice result.
\begin{theorem}
    Let $x^{*}$ be local minimum of \ref{eq: gen_constraned}, fulfilling
    Abadie CQ. Then there exist Lagrange multipliers $\lambda^{*} \in
    \mathbb{R}^{m} $ and $\mu \in \mathbb{R}^{p}$, such that $(x^{*},
    \lambda^{*},\mu^{*})$ is KKT point of the constrained optimization
    problem.
\end{theorem}
\begin{proof}
   The strategy is to utilize the cone duality and already known results.
   From Proposition \ref{prop: descent} we know that $\nabla f(x^{*}) \in
   (T_X(x^{*}))^{*} $, and since $x^{*}$ fulfills Abadie CQ it holds that
   $(T_X(x^{*}))^{*}  = (T_{\text{lin}}(x^{*}))^{*}$. Thus $-\nabla f(x^{*})
   \in -(T_{\text{lin}}(x^{*}))^{*}$. From Theorem \ref{thm: duality} we have
   that $-\nabla f(x^{*}) \in N_{\text{lin} }(x^{*})$ and hence we can write
   \begin{align}
       -\nabla f(x^{*}) = \sum_{i \in \mathcal{A}(x^{*}) } \lambda_i^{*}  \nabla g_i(x^{*}) +
    \sum_{j=1}^{p} \mu_j^{*}h_j(x^{*}),
   \end{align}
   where we know there are $\lambda^{*}_i \ge 0$ for all $i \in
   \mathcal{A}(x^{*})$ and $\lambda^{*}_i = 0$ for all $i \in
   \mathcal{I}(x^{*})$ and $\mu^{*}_j \in \mathbb{R}$ for all $j=1,\ldots,p$.
   Then we expand the sum and rearrange to get
   \begin{align}
       &\nabla f(x^{*}) + \sum_{i=1}^{m} \lambda^{*}_m\nabla g(x^{*}
    + \sum_{j=1}^{p} \mu^{*}_j \nabla h_j(x^{*}) = 0
   \end{align}
   which is the first condition in the KKT conditions
   \begin{align}
       \nabla_x L(x^{*}, \lambda^{*}, \mu^{*}) = 0.
   \end{align}
   Also we have that if $i$ is active $g_i(x^{*}) = 0$ and if $i$ is inactive
   we have $\lambda_i^{*}=0$. Thus  $\lambda^{*}_i \ge 0$, $g_i(x^{*})\le 0$,  $\lambda_i
g(x^{*}) \le 0$. Also by construction of the problem $h_j(x)=0$ for all
$j=1,\ldots,p$.
\end{proof}
\section{Convex Optimization Problems}
Since Abadie CQ is an excellent result since it gives a hint how to find
$x^{*}$. But it assumes something about $x^{*}$ before it is even found. In
the case of convex optimization problems we can derive a much relaxed
sufficient condition which implies Abadie CQ. Firstly a brief definition of
convex functions
\begin{definition}
    Let $U \subseteq \mathbb{R}^{n}$, the function $f: U \to \mathbb{R}$
    is called convex if for all $x, y \in U$ and $\lambda \in [0, 1]$ it
    holds that
    \begin{align}
        f(\lambda x + (1-\lambda)y) \le \lambda f(x) + (1-\lambda)f(y).
    \end{align}
\end{definition}
An important result we will use throughout is the following.
\begin{proposition}
    \label{prop: convexity}
    Let $U \subseteq \mathbb{R}^{n}$ be a nonempty, open and convex set, $f:
    U \to \mathbb{R}$ a differentiable function on $U$. Then the following
    statements are equivalent
    \begin{enumerate}
        \item $f$ is convex on $U$;
        \item $\nabla f(x)^{T}(y-x) \le f(y)- f(x)$ for all $x, y \in U$;
        \item $(\nabla f(x) - \nabla f(y))^{T}(y-x)\ge 0$ for all $x, y \in
            U$;
        \item if $f \in C^{2}(U)$, then $\Delta f(x)$ is positive
            semidefinite for all $x \in U$.
    \end{enumerate}
\end{proposition}
\begin{proof}
    We show (1) $\Leftrightarrow$ (2) $\Leftrightarrow$ (3) and (2)
    $\Leftrightarrow$ (4). Starting off with (1) $\Rightarrow$ (2) by definition
    we have that $\forall x, y \in U$ and $\lambda \in [0, 1]$ it holds that
    \begin{align}
        f(\lambda x + (1-\lambda)y) \le \lambda f(x) + (1-\lambda)f(y).
    \end{align}
    Rearranging, dividing by $\lambda$ and expanding the fraction by $x-y$ we
    get
    \begin{align}
        \frac{f(y+\lambda(x-y)) - f(y)}{\lambda (x-y)}(x-y) \le f(x) - f(y)
    \end{align}
    Letting $\lambda \downarrow 0$ we get
    \begin{align}
        \nabla f(y)^{T} (x-y) \le f(x) - f(y).
    \end{align}
For (2) $\Rightarrow$ (1), we know that $U$ is convex hence fix $x, y \in U$
then for any $z$ in the line segment of $x$ and $y$ there is a $\lambda \in
[0,1]$ such that $z = \lambda x + (1-\lambda)y$. We can construct two
inequalities
\begin{align}
    &\nabla f(z)^{T} (x-z) \le f(x) - f(z) \\
    &\nabla f(z)^{T} (y-z) \le f(y) - f(z).
\end{align}
By multiplying the first one with $\lambda$ and second one with $(1-\lambda)$
and adding them together we get
\begin{align}
    \nabla f(z)^{T} (\lambda x + (1-\lambda)y - z) \le \lambda f(x) +
    (1-\lambda)f(y) - f(z).
\end{align}
Inserting for $z$ we get $\lambda x+ (1-\lambda)y - z =0$ and thus
\begin{align}
    f(z) = f(\lambda x + (1-\lambda)y) \le \lambda f(x) + (1-\lambda)f(y).
\end{align}
For (2) $\Rightarrow$ (3) consider two inequalities for all $x, y \in U$
\begin{align}
    \nabla f(x)^{T} (y-x) \le f(y) - f(x) \\
    \nabla f(y)^{T} (x-y) \le f(x) - f(y).
\end{align}
Adding them together gives
\begin{align}
    (\nabla f(x) - \nabla f(y))^{T} (x-y) \le 0 .
\end{align}
For (3) $\Rightarrow$ (2) consider a function $g(t) = f(t x + (1-t)y)$
defined on the line segment of $x, y \in U$ for $t \in [0, 1]$ and
well-defined since $U$ is convex. The derivative is
\begin{align}
    g'(t) = \nabla f(t x + (1-t) y) (y-x)
\end{align}
Note that by (3) we have that
\begin{align}
    \nabla f(x)^{T} (y-x) \le \nabla f(y)^{T} (y-x),
\end{align}
for all $x, y \in U$. Now consider the Mean Value theorem for $g$,
\begin{align}
    f(y) - f(x) &= \int_0^1 g'(t) dt\\
                &= \int_0^1 \nabla f(tx +(1-t)y)(y-x)dt\\
                &\ge \int_0^1 \nabla f(x)(y-x) dt\\
                &= \nabla f(x) (y-x).
\end{align}
Lastly for (2) $\Leftrightarrow$ (4) use Taylor's Theorem, i.e. for all $x, y
\in X$ we have
\begin{align}
    f(y) = f(x) + \nabla f(x)^{T} (y-x) + \frac{1}{2}(y-x)^{T} \nabla^2
    f(\xi)(y-x),
\end{align}
for some $\xi$ in the line segment of $x$ and $y$. For $\Rightarrow$ we have
that the Hessian is positive semidefinite $\frac{1}{2}(y-x)^{T} \nabla^2 f(\xi)(y-x) \ge 0$
thus rearranging gives
\begin{align}
    \nabla f(x)^{T} (y-x) \le f(y) - f(x)
\end{align}
On the other hand we have that $f$ is convex, i.e. $0 \le f(y)-f(x)- \nabla
f(x)^{T} (y-x) $ and thus
\begin{align}
    (y-x)^{T} \nabla^2 f(\xi)(y-x) \ge 0.
\end{align}
Since $x, y$ was arbitrary and $\xi$ in the line segment of $x$ and $y$ we
can conclude that the Hessian is positive semidefinite which concludes the
whole proof.
\end{proof}
Throughout this section consider the following convex problem
\begin{align}
    \min\;&f(x) \label{eq: convex_constrained}\\
    \text{s.t.:}\;&g_i(x) \le 0,\; i=1,\ldots,m \nonumber\\
        &Ax = b,\; \nonumber\\
        &x \in \mathbb{R}^{n}\nonumber,
\end{align}
where $f, g_i$ are convex and continuously differentiable functions for all
$i=1,\ldots,p$ and the function $h$ becomes affine linear, i.e. $h(x) =
Ax-b$, where $A \in \mathbb{R}^{p\times n}, b\in \mathbb{R}^{p}$. The
feasible set $X$ of problem \ref{eq: convex_constrained} is a convex set in
this case. An important condition for convex problems is the so called Slater
constraint qualification (CQ).
\begin{definition}
    The optimization problem \ref{eq: convex_constrained} fulfills the Slater
    CQ if there exists an $x' \in \mathbb{R}^{n}$ such that
    \begin{align}
        &g_i(x') < 0,\; i=1,\ldots,m \\
        &Ax' = b.
    \end{align}
\end{definition}
Given Slater CQ we can prove the following

\begin{theorem}
    Let $ x^{*} $ be a local minimum of \ref{eq: convex_constrained}, which
    fulfills Slater CQ. Then there exist Lagrange multipliers $\lambda^{*}
    \in \mathbb{R}^{m}$ and $\mu^{*} \in \mathbb{R}^{p}$  such
    $(x^{*},\lambda^{*},\mu^{*})$ is a KKT point of \ref{eq:
    convex_constrained}.
\end{theorem}
\begin{proof}
We show that $x^{*}$ fulfills Abadie CQ, i.e. $T_{\text{lin}}(x^{*})
\subseteq T_X(x^{*})$. Let $d \in T_{\text{lin}}(x^{*})$ and choose an $x'
\in X$ satisfying Slater CQ. Set $d' := x' - x^{*}$, by Proposition
\ref{prop: convexity} (2) we have for all $i \in \mathcal{A}(x^{*})$ that
\begin{align}
    \nabla g_i(x^{*})^{T} d'
    &= \nabla g_i(x^{*})^{T} (x^{*} + d' - x^{*} )\\
    &= g_i(x^{*} +d') -g_i(x^{*})

\end{align}
\end{proof}
Thus Salter CQ implies Abadie CQ and we can show that KKT optimality
conditions are sufficient for optimality
\begin{theorem}
    Let $(x^{*}, \lambda^{*},\mu^{*})$ be a KKT point of
    \ref{eq: convex_constrained}. Then $x^{*}$ is a local minimum of
    \ref{eq: convex_constrained}.
\end{theorem}
\begin{proof}
TODO
\end{proof}


