\chapter{Introduction}
Maximal Extractable Value (MEV), sometimes also called Miner Extractable
Value or Block Proposer Extractable Value, refers to the maximal value that
can be extracted from block production (excluding block reward and gas fees)
by including, excluding and changing the order of transactions in a block and
thereby creating an economic gain. The opportunities, among others, include
arbitrage, frontrunning and liquidations. By definition, MEV exists on any
decentralized blockchain. This thesis primarily focuses on the
application found on the Ethereum blockchain. Nevertheless, everything can be
applied to other blockchains with smart contract functionality. Maximizing
MEV extraction requires sophisticated technicalities, custom software and
high-end hardware, making it more likely for institutional centralized
entities to outperform individuals. The centralization is a threat to the
most important quality of blockchain technology, decentralization. There are
multiple proposals on how to combat MEV centralization, however solutions that
evolve completely removing MEV from the picture close one door but open
multiple other ones.
\newline
Instead, the Ethereum Foundation \cite{ethereum_foundation} and Flashbots
\cite{flashbots} implemented a reshape of `MEV economics' through the
Proposer-Builder-Separation (PBS) \cite{proposer_builder_sep}. In short the
PBS splits the MEV reward through multiple entities, each playing a distinct
role in the ecosystem. The Searchers capture the bulk of the MEV profits.
They identify the opportunities and commit bundle transactions to exploit
them. The Block Builders construct the blocks by including, excluding and
optimizing a set of selected transactions to produce the most profitable
blocks. They earn a share of MEV from the Searchers, who are committing bids
to the Block Builders to include the bundles. Block Builders can also earn
MEV directly by reordering transactions. The last entity that takes a direct
cut of MEV are Validators. They propose the most profitable block to the
network and outsource block production to block builders via software like
MEV-Boost \cite{flashbots}. They earn a share of MEV by competing bids of
Block Builders to include their block. In multiple stages the PBS produces a
highly competitive environment where usually the highest bid wins.
Nevertheless, PBS is not a perfect solution. There have been cases where, in
one month, $80\%$ of the blocks have been produced by only two block
builders and the majority of the opportunities found by a handful of
Searchers.
\newline
In the scope of the Ethereum blockchain, this thesis is primarily applicable
for Searchers. The aim of the thesis is to contribute to the decentralization
of MEV by `laying it out in the open'.
\newline
We focus on a specific opportunity arising from an emerging form of financial
markets. These are implemented on Decentralized Exchanges (DEXs). In
comparison to traditional markets, based on order books and centralized
liquidity market makers, the DEXs are based on Automated Market Makers (AMMs)
and decentralized liquidity providers. AMMs, specifically Constant Function
Market Makers (CFMMs), use a trading function and set of rules to
automatically enable trading and liquidity provision for any user satisfying
the rules. The opportunity in question in this thesis is an $n$-cyclic
arbitrage over multiple DEXs, in which an optimized sequence of trades on $n$
distinct DEXs is executed. The sequence starts from one specific asset,
executes the trades, and ends on that same asset with a balance greater than
before the trades. We model CFMMs to construct an optimization problem that
maximizes the profit of an $n$-cyclic arbitrage. In this scenario, Searchers
races against the block time, trying to solve multiple of these optimization
problems to find a profitable one. When the clock turns and a new block
is included, the market equilibrium changes and the optimization problem
becomes obsolete. One of the improvements in this case is to make a convex
optimization problem. The original implementation allows for a convex
relaxation. We can show that the relaxation has the same optimum as the
original. This allows us to utilize more efficient convex solvers.
\newline
There are two further problems with the proposed optimization problem. These
can be solved through one solution. The current model optimizes a trade
vector over a set of DEXs, yet there are currently around $4 \cdot 10^{5}$
DEXs deployed on UniswapV2 and UniswapV3 \cite{uniswap} alone, not counting
other implementations. It would be extremely inefficient to optimize over all
of these DEXs, and it would still be inefficient and unprofitable to optimize
on a restriction of a much smaller set of DEXs. The second problem is that the
optimal trade vector does not specify in which order the trades should be
executed. We propose a solution to these problems by finding a specific walk
on a highly connected graph of DEXs. The graph has nodes representing assets
traded on the DEXs. These nodes are connected by an edge if the underlying
assets can be traded through a DEX. The weights of the edges are
the so-called marginal exchange rates. The goal is to find a walk starting
and ending at the same node, maximizing the product of the marginal exchange
rates without repeating an edge. The information of this optimal walk can
then be used to construct the convex optimization problem and execute the
trade by the chronology of the walk. Any brute-force,
depth/breadth-first-search solution would take too long. Hence, we propose a
Markov Decision Process representing the walk optimization. We find an
optimal policy through a modified policy gradient method, specifically a
variation of the Monte Carlo Tree Search REINFORCE \cite{sutton_barto}
algorithm inspired by AlphaZero \cite{alphazero}.
\newline
The chapter outline of the thesis is the following.
\newline
In the second chapter we go through the background on first-order optimality
conditions for constrained nonlinear optimization problems based on
\cite{bot_optimization}. Through the characterization of optimal solutions
based on tangent cones, we can show that an optimal solution, under certain
constraint qualifications, implies the existence of the Lagrange multipliers
satisfying the Karush-Kuhn-Tucker (KKT) conditions. These conditions are
necessary for optimality. In the case where the objective function and the
inequality restrictions are convex and the equality restrictions are affine linear,
we can apply the Slater constraint qualification. Under this assumption, the
KKT conditions are both necessary and sufficient for optimality.
\newline
In the third chapter, we model DEXs through CFMMs. By gradually building the
foundations of general trades, we can construct an optimization problem
optimizing a trade over one CFMM. We can show that the convex relaxation of
the problem, where we relax the equality constraints to inequality
constraints, has the same optimal solution as the original. We can
inductively increase the number of CFMMs for which we optimize a trade
vector, giving a general optimization problem for a $n$-cyclic arbitrage.
\newline
In the fourth chapter, we give a brief introduction to Markov Decision
Processes (MDPs), Graph Neural Networks (GNNs), Monte Carlo Tree Search
(MCTS) and a policy gradient method called REINFORCE. All of these
will be utilized for construction of the algorithm used to train the policy
parametrized by a deep graph neural network for finding an $n$-cyclic
arbitrage walk.
\newline
In the fifth and final chapter, we model the underlying MDP of finding the
$n$-cyclic arbitrage walk. We describe the specific architecture of GNN
parametrizing the policy and the implementation of a modified REINFORCE MCTS
parameter updates. Lastly, we give an overview of the results of the policy
optimization.
