\include{./preamble.tex}

\begin{document}

\maketitle

\section{Automated Market Makers}
Decentralized finance (DeFI) consist of automated financial tools providing the
ability, through smart contracts on the blockchain, to lend or borrow funds,
trade crypto currencies and among others provide market liquidity earning
an interest. The intermediary mechanism are Automated Market
Makers (AMMs), which are governed by the Constant Function Market Maker
(CFMM) trading function. CFMM describes the relation between the liquidity of
two or more assets, but for simplicity sake consider for further on that we
are only looking at a liquidity pools of only two assets. There are two
options when interacting with a CFMM: providing liquidity and trading one
asset for the other one or vice versa. Users who are providing liquidity are
called liquidity providers (LPs) and users who are trading (or taking
liquidity) are called liquidity takers (LTs). The interactions of these users
with the CFMM are governed by two rules, the LT trading condition and the LP provision
condition.
\newline

CFMM allows us to value two assets X and Y against each other, by initially
providing a quantity $x \in \mathbb{R}_+$ of asset X and a quantity $y \in
\mathbb{R}_{+}$ of asset Y as liquidity. The pair $(x,y)$ is called the
reserves of the liquidity pool. The CFMM is a trading function of the
reserves $f: \mathbb{R}_+^{2} \to \mathbb{R}_+$. A user is willing to
exchange $\delta x$ of asset X through the CFMM. The amount $\delta
y$ of asset $Y$ the user is going to receive is governed by the LT trading
condition which requires the trading function to be constant throughout the
trade. After the trade is executed the new reserves are updated to
$(x+\delta x, y - \delta y)$, the amount $\delta y$ the user
receives is determined implicitly through
\begin{align}
    f(x + \delta x, y - \delta y) = f(x, y) = k, \label{eq: 1}
\end{align}
where $k \in \mathbb{R}_+$ is a constant. The LT trading condition
implies a level function $\phi_k(x) = y$, which is just $f(x,y) = k$
rewritten. The user exchanging $\delta x$ will receive $\delta y = \delta x
\cdot \tilde{Z}(\delta x)$, where $\tilde{Z}(\delta x)$ is the exchange rate
of asset Y wrt. asset X. Rearranging gives
\begin{align}
    y - \delta y = \phi_k(x + \delta x) \implies \phi_k(x) - \delta
    x\tilde{Z}(\delta x) = \phi (y + \delta y),
\end{align}
thereby the exchange rate is
\begin{align}
    \tilde{Z}(\delta x) = \frac{\phi_k(x) - \phi_k(x+\delta x)}{\delta x}.
\end{align}
We can also define the marginal exchange rate for a very small trade $\delta x$
\begin{align}
    Z = \lim_{\delta x \to 0} \frac{\phi_k(x) - \phi_k(x+\delta x)}{\delta x}
    = -\phi_k'(\delta x), \label{eq: marginal exchange rate}
\end{align}
which can be taught of as the exchange rate without any execution cost. The
execution cost is referred to as $|Z - \tilde{Z}(\delta x)|$.
\newline
Further more the condition $\delta x \cdot \tilde{Z}(\delta x) = \delta y$ defines an
exchange function $F: \mathbb{R}_+ \to [0, y)$, where $F(\delta x) =
\delta y$ bijectively satisfying the condition of the $\delta y$ for a given
$\delta x$ in \ref{eq: 1}. Vice versa we can define also a reverse exchange
function $G : [0, y) \to \mathbb{R}_+$, satisfying $G(\delta y) = \delta
x$. These function are inverses of each other $F(G(\delta y)) = \delta x$ and
$G(F(\delta x)) = \delta y$. Both functions can be solved for numerically
through the relation in \ref{eq: 1}. The domain/image $[0, y)$ is a
restriction, such that the reserves of the liquidity pool are not completely
drained.
\newline

On the other hand the LP provision condition is the rule determining the LP
transaction, where the user deposits or withdraws quantities $(\delta x,
\delta y)$ of asset X and Y respectively from the liquidity pool. LP
provision condition states that the marginal exchange rate $Z$ does not
change before and after the transaction is executed. Let $k_1$ and $k_2$ be
the depth constant of the pool before and after the transaction, satisfying
\begin{align}
    f(x, y) = k_1, \qquad f(x+\delta x, y+\delta y) = k_2,
\end{align}
with corresponding level functions $\phi_{k_1}$ and $\phi_{k_2}$. From the
definition of the marginal exchange rate \ref{eq: marginal exchange rate} we
deduce the LP provision condition to
\begin{align}
    -\phi'_{k_1}(x) = -\phi'_{k_2}(x + \delta x)
\end{align}
\newline

The most commonly adopted CFMM is the Constant Product Market Maker. The
trading function is $f(x, y) = xy$, the level function $\phi_k(x) =
\frac{k}{x}$, the marginal exchange rate $Z = \frac{y}{x}$ and the forward
and the reverse exchange functions are
\begin{align}
    &F(\delta x) = y\left(1-\frac{x}{x+\gamma \delta x}\right),\\
    &G(\delta y) = \frac{x}{\gamma}\left(\frac{y}{y-\delta y} - 1\right),
\end{align}
where $(1-\gamma) \in (0,1)$ is the fee for the exchange associated with the
pool. Usually this fee is $0.3\%$ but can vary from pool to pool. Trading
$\delta x$ for $\delta y$, the pool would charge a $1-\gamma$ fee, essentially
trading only $\gamma \delta x$ for $\delta y$ or vice versa. The fee is then
split by liquidity providers, contributing to their reserves. Also notice
that the forward exchange rate is convex and the reverse exchange rate is
concave in this example, it is possible to prove this holds in general.
\section{Frontrun Attack}
A frontrun or sandwich attack is an exploit of a trade with large price
impact on a AMM on the blockchain. It occurs when a victim commits a
transaction simply trading $\delta x_V$ (a large amount wrt. the reserves) of
asset X for $\delta y$ of asset Y on a pool with reserves $(x, y)$. Once the
transaction is committed, but not yet executed by the block builders it lands
in the transaction pool. The attacker identifies the transaction before it is
executed and decides commit two transactions, sandwiching the victims
transactions. The first transaction buys asset Y for asset X before the
victim and the second transaction buys asset X for asset Y, occurring
immediately after the victims trade. Decentralized exchanges allow the user
to set a slippage limit $\eta \in (0, 1)$ which indicates how much of a price
impact they are willing to tolerate before their transaction is executed. If
the price impact is bigger than the indicated slippage limit the transaction
will be reverted and the trade will not be executed. The
aim is to maximize the attackers first trade amount wrt. the victims
slippage limit.
\subsection{Without fees}
Modeling this problem without the $(1-\gamma) \in (0,1)$ fee collected for
each trade buy the pool is simpler because the reserves are updated without
having to incorporate the fee. The victim buys an amount $\delta x_V$ of
asset X, buy the forward exchange function the user receives $\delta y =
F(\delta x_V)$ for asset Y and is willing to tolerate only receiving
$(1-\eta)F(\delta x_V)$. On the other hand the attacker executes a trade of
$\delta x$ receiving $F(\delta x)$. The amount of asset Y received by the
victim after the attackers trade is executed is $F(\delta x + \delta x_V) -
F(\delta x)$. The victims trade will only be executed if
\begin{align}
    F(\delta x_V + \delta x) - F(\delta x) \le (1-\eta) F(\delta x_V).
\end{align}
Then we have the following optimization problem.
\begin{align}
    \max \quad  &\delta x \\
    \text{s.t.:}\quad & F(\delta x_V + \delta x) - F(\delta x) \le (1-\eta)
    F(\delta x_V) \\
              & \delta x - x \le 0 \\
              & F(\delta x) - y \le 0 \\
              & \delta x > 0.
\end{align}
\subsection{With fees}
With the pool fee the LT trading condition is $f(x+\gamma \delta x, y - \delta y) =
k$, after the trade the reserves are updated to $(x, y) \mapsto (x + \delta x, y - \delta
y)$. For the attackers trade the LT trading condition is governed by the
initial depth constant of the pool $k_0$, meaning $f\left(x+\gamma \delta x,
y - \delta y\right) = k_0$, after which the reserves are $(x', y') =
(x+\delta x, y+\delta y)$. Following the attackers trade, the victims trade
of $\delta x_V$ is executed under the condition $f(x' + \gamma \delta x_V, y'
- \delta y_v) = k_1$. Where $k_1$ is the depth constant of the pool after the
attackers trade. This means that the forward exchange rate at $k_0$ and $k_1$
is different and dependent on the fee $(1-\gamma)$. Let us denote $F_{k_0}:
\mathbb{R}_+ \to \mathbb{R}$ the forward exchange rate at $k_0$ and $F_{k_1}:
\mathbb{R}_+ \times \mathbb{R}_+ \to \mathbb{R}$  the forward exchange rate
at $k_1$. The forward exchange rate $F_{k_1}$ is a function of the victims
input amount and the attackers input amount, which tells us how the reserves
are updated, i.e. $F_{k_1}(\delta x_V, \delta x) = \delta y_V$. Under the slippage
limit of the victim $\eta \in (0,1)$ the condition that the victims trade is
executed is then
\begin{align}
    F_{k_1}(\delta x_V, \delta x) \le (1-\eta)F_{k_0}(\delta x_V).
\end{align}
This is the bound for the optimization problem. As an example for the CFMM
$f(x,y) = k$, $F_{k_0}$ and $F_{k_1}$ are
\begin{align}
    &F_{k_0}(\delta x_V) = y \left(1-\frac{x}{x+\gamma \delta x_V}\right) \\
    &F_{k_1}(\delta x_V, \delta x) = \left(y - F_{k_0}(\delta x)\right)
    \left(1-\frac{x+\delta x}{x + \delta x +\gamma \delta x_V}\right)
\end{align}
\section{Arbitrage}
The arbitrage opportunity/attack utilizes the lack of market price
equilibrium of an asset to preform a sequence of trades on different AMMs to
achieve a risk free profit. There are two challenges concerning this problem,
first is finding the right sequence of AMMs on which the trades will be
executed , and the second is to find the right trade amount that will
maximize the profit. After the arbitrage is executed the market price of the
asset wrt. the AMMs will be in an equilibrium.
\subsection{Optimization Problem}
First simply consider a two pool arbitrage (Pool 1, Pool 2) with the same
CFMM trading function and same fee $(1-\gamma)\in (0,1)$, where two pools are
both exchanging asset X for asset Y or vice versa. The reserves and the depth
constants of the pools are $(x_1, y_1)$ and $(x_2, y_2)$, $k_1$ and $k_2$
respectively. The marginal exchange rates for asset Y in terms of asset X in
the two pools are $Z_1$ and $Z_2$. There is an arbitrage opportunity to trade
$\delta x$ amount of asset X in Pool 1 for $\delta y$ amount of asset Y,
trade $\delta y$ of asset Y in Pool 2 for $\delta x'$ of asset X, when $Z_1 <
Z_2$. The equilibrium is achieved when $Z_1 = Z_2$. The goal is to maximize
the profit $\delta x' - \delta x$ wrt. the LT trading conditions on both
pools. The LT trading condition for the first trade is $f(x_1+\gamma \delta
x, y_1 - \delta y) = k_1$ and for the second trade $f(x_2 - \delta x',
y_2 + \gamma \delta y) = k_2$. Here $\delta y$ and $\delta x'$ can be calculated with the
forward and reverse exchange functions, dependent on $\delta x$. The profit
optimization then is
\begin{align}
    \max \quad &\delta x' - \delta x \\
    \text{s.t.:}\quad &f(x_1+\gamma \delta x, y_1 - \delta y) = k_1\\
             &f(x_2 - \delta x', y_2 + \gamma \delta y) = k_2\\
             &\delta x > 0.
\end{align}
Generally the trading function is not affine linear hence this optimization
problem is not convex. But in the case that the trading function is convex
the relaxation of the problem, changing $=$ to $\le$ is a convex optimization
problem and can be solved efficiently. Alternatively we could possibly
reformulate the constraints wrt. the forward and reverse exchange function
which are concave/convex functions and arrive together with a relaxation to a
convex optimization problem. We would need to check if the relaxation is then
at least quasi optimally solving the same problem.

The general case of an arbitrage going through multiple AMMs is a circuit path
\begin{align}
    X_1 \xrightarrow{\text{Pool 1}} X_2 \xrightarrow{\text{Pool 2}}
    \hdots \xrightarrow{\text{Pool n-1}} X_n \xrightarrow{\text{Pool n}} X_1,
    \label{eq: path}
\end{align}
where the assets $\{X_1, \ldots, X_n\}$ are not necessarily unique, but the
pools $\{\text{Pool 1},\ldots,\text{Pool n}\}$ are unique. Assume that the
pools can only trade two assets. For all $i \in \{1,\ldots,n\}$ each Pool $i$
allows trading of assets $X_i$ and $X_{i+1}$ with a fee of $(1-\gamma_i) \in
(0,1)$. It has a LT trading condition with a CFMM $f_i:\mathbb{R}_+^{2}\to
\mathbb{R}$, as arguments the reserves of the pool $x_i = (x_{i_1}, x_{i_2})$ and
the initial depth constant of the pool is $f_i(x_{i_1}, x_{i_2}) = k_i$. The
level function of the pool is $\phi^{i}_{k_i}(x_{i_2}) = x_{i_1}$ and its
marginal exchange rate is $Z_i = -(\phi^{i}_{k_i})'(x_{i_2})$. The marginal
exchange rate of Pool i, $Z_i$ is the marginal exchange rate of exchanging
the asset associated with the reserve $x_{i_1}$ for the asset associated with
the reserve $x_{i_2}$. W.l.o.g assume that arguments of the trading function,
i.e. reserves $(x_{i_1}, x_{i_2})$, are sorted in the sense where a trade
allways consists of exchanging an amount of asset associated with reserve $x_{i_1}$ for
an amount of an asset associated with the reserve $x_{i_2}$.
The arbitrage trade is sequentially conducting trades through all pools from
asset $X_1$ back to asset $X_1$. In each pool a trade is executed set by a
trade vector $\xi_i = (\xi_{i_1}, \xi_{i_2}) \in \mathbb{R}^{2}$, exchanging $\xi_{i_1}>0$ for
$\xi_{i_2}<0$. At each pool the trade is bound by the LT trading condition
$f_i(x_i+\xi_i) = f_i(x_i) = k_i$. The goal is to find the trade vectors
$\xi_i$ that maximize the profit $\xi_{n_1} - \xi_{1_0}$, putting the market
price at the equilibrium. The modeling leads us to the following optimization
problem
\begin{align}
    \max \quad &\xi_{n_1} - \xi_{1_0} \\
    \text{s.t.:}\quad &f_i(x_i+\xi_i) = f_i(x_i) \quad \quad \forall\;
    i \in \{1,\ldots,n\}\\
             &\xi_{1_0} > 0.
\end{align}
\subsection{Circuit Routing Problem}
The challenging part in the arbitrage problem is finding the opportunity,
i.e. finding the path in \ref{eq: path} in an extremely short amount of time.
If thought of as a graph problem, the set of vertices $V$ are assets $\{X_1,
\ldots, X_n\}$, and edges connecting them are pools $E = \{\text{Pool 1},
\ldots, \text{Pool m}\}$, where each pool $k$ is represented by an edge
$e_{ij}^{k}$ connecting asset $X_i$ with asset $X_j$. Each edge has a weight
associated to it $w^{k}_{ij}$, which we set as
\begin{align}
    w_{ij}^{k} = -\log(Z_{ij}^{k}),
\end{align}
where $Z_{ij}^{k}$ is the marginal exchange rate of asset $X_j$ in reference
to asset $X_i$ at pool $k$. Additionally note that the graph is a multigraph:
a vertex can have multiple edges, one asset can share multiple pools. The aim
is to find a path $C\subseteq E$ starting of an arbitrary length $l \le m$
which starts and ends at a vertex $X_m$, $m\in\{1,\ldots,n\}$ with non
repeating edges, which minimizes the weights. Basically a circuit minimal
weight problem.
\newline

To solve this problem given the time constraints we can
resort to using Monte-Carlo-Tree-Search (MCTS) based reinforcement learning.
The first steps are to give a mathematical optimization model of the problem
and translate it to a Markov-Decision-Process (MDP). Then create an
appropriate Graph Neural Network (GNN), a message passing GNN (GMPPN) with
appropriate encoders. Finally train the network through two pipelines, first the
supervised learning pipeline which passes the parameters into reinforcement
learning and MCTS pipeline, retraining the reward function.
\newline

The mathematical model of the problem is
\begin{align}
    \min \quad& \sum_{(i,j,k) \in E} \delta_{ij}^{k} w_{ij}^{k} \label{eq:
    obj function}\\
    \text{s.t.:}\quad & \delta_{ij}^{k} + \delta_{ji}^{k} \le 1 \quad \forall
    (i,j,k) \in E, \label{eq: constraint 1} \\
    & \sum_{j\in V, j\neq i} \delta^{k}_{ij} = \sum_{j \in V, j\neq i}
    \delta^{k}_{ji} \quad \forall i \in V \label{eq: constraint 2} \\
    & \text{3rd constarint: possibly subcycle elimination in the tour?}
\end{align}
where $\delta_{ij}^{k}$ is the binary decision variable, if $(i,j,k) \in C$
then $\delta_{ij}^{k} = 1$ and 0 otherwise. Where \ref{eq: obj function} is
the objective function, constraint in \ref{eq: constraint 1} esures that the
edge $(i, j, k)$ is traversed only once. The constraint in \ref{eq:
constraint 2} ensures flow conservation, for any vertex the number of
incoming and outgoing edges must be equal.
\newline

As for the next steps I will have to get more familiar with MDPs, MCTS, GNNs...

\section{Just in Time Liquidity}
Just in Time Liquidity (JIT Liquidity), is an attack, where an attacker is
also frontrunning a victims transaction. On contrast to the frontrun sandwich
attack, the attacker actually gives the victim a worse exchange rate. Whereas
JIT Liquidity actually provides the victim with a better exchange rate. In
essence the attacker will execute a LP trade, lowering the victims exchange
rate and immediately afterwards pull liquidity with another LP trade
collecting the victims trading fees. While this form of attack can be done on
AMMs described above, it is usually not profitable because all LP trades add
liquidity on the whole price range, hence splitting the trading fees equally
on all liquidity providers. The JIT liquidity attack is usually done on AMMs
where the liquidity providers are able to provider liquidity on a specific
price range $(Z_{i}, Z_{i+1}]$.
\subsection{Concentrated Liquidity Markets}
We will focus here on liquidity provision on CPMMs with concentrated
liquidity (CL). In AMMs (pools) with CL the space of exchange rates (or only
rates) where the LP provides liquidity is split into a set of values
$\{Z_i\}_{i \in \mathbb{Z}}$, which are called ticks. Commonly used ticks are
$Z_i = 1.0001^{i}$. Each tick range $(Z_i, Z_{i+1}]$ has an associated depth
constant, which determines the LP trading condition when the marginal
exchange rate $Z$ is in this interval.

\section{TODO}
The following are the obvious next steps:
\begin{itemize}
    \item Optimally conditions to the optimization problems
    \item Solutions to the optimization problems
    \item Equilibrium condition for general arbitrage with n pools
    \item Derive MDP for circuit problem
    \item MCTS reinforcement learning for the circuit problem (big)
    \item JIT Liquidity further formalization, together with concentrated
        liquidity AMMs and optimization problem.
\end{itemize}

\end{document}
